Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-05-17 scripts/MouseSexDiffApril20_2020.Rmd
Updating supplement

---
title: "Sex and Power: sexual dimorphism in trait variability and its eco-evolutionary and statistical implications"
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgortaza Lagisz, Jeremy Mason, Daniel Noble, Alistair
  Senior & Shinichi Nakagawa
date: "April 2020"
output:
  html_document:
    code_download: yes
    code_folding: hide
    depth: 4
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    toc: yes
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

# Set-up
## 1. Packages

@@

```{r packages}
library(pacman)
pacman::p_load(readr, dplyr,metafor, devtools, purrr, tidyverse, tidyr, tibble, kableExtra, robumeta, ggpubr, ggplot2, png, grid, here)
```

## 2. Functions 
Functions needed for preparing the data for meta analyses

### 1) Sub-setting data
Create a function for sub-setting the data to choose only one data point per individual per trait: "data_subset_parameterid_individual_by_age"

@@

### 2) Population statistics
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

### 3) Extraction of effect sizes and sample variances
Function: "create_meta_analysis_effect_sizes"

@@

## 3. Load & clean data

### 1) Data loading and cleaning of the csv file

@@

# Meta-analyses
## 1. Population as analysis unit 
(Step C, Figure 3 in main document)

### 1.1: Loop: Meta-analyses on all traits

@@

### 1.2: Merging datasets & removal of non-converged traits

@@

## 2. Meta-analysis: condensing non-independent traits 
(Step F in Figure 3 in main article)

### 2.1: Dealing with Correlated Parameters, preparation

@@

### 2.2: Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

### 2.3: Meta-analyses on correlated (sub-)traits, using `robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

## 3. Second-order meta-analysis for functional groups
(Section H in Figure 3 in main article)

### 3.1: Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR
#### 3.1.1: Preparing the data

@@

#### 3.1.2: Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

# Visualising meta-analysis results
## 1. Preparation for Figure 4 
#### 1.1: Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

#### 1.2: Overall results of second order meta-analysis (Figure 4, Panel B)
#### 1.3: Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

## 2. Figure 4

@@

## 3. Preparation for Figure 5
#### 3.1: Preparing data for traits with CI not overlapping 0
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

## 4. Figure 5
Sex biases in trait values across trait groups (Step J in Figure 3). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

# Supplemental Tables and Figures

## Supplemental Tables
#### 1. Supplemental Table S1: "Strains and Center Sample Sizes"
  This table summarises the available numbers of male and female mice from each strain and originating institution.

```{r Table1_prep}
  # Initial checks
  length(unique(data$parameter_name)) # 232 traits
  length(unique(data$parameter_group)) # 161 parameter groups
  length(unique(data$procedure_name)) # 26 procedure groups
  length(unique(data$biological_sample_id)) # 27147 individial mice  
  # To show a table with number of males and females per strain per production center:

  ST1 <- (cbind(data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf))) 
```
```{r Table1}

  kable(ST1) %>%
    kable_styling() %>%
    scroll_box(width = "60%", height = "200px")

#### 2. Supplemental Table S2: "Numbers of correlated and uncorrelated traits"

```{r, TableS2}  
# Table detailing numbers of correlated and uncorrelated traits
kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "60%", height = "200px")
```

#### 3. Supplemental Table S3: "Meta-analysis results"

  We use this corrected (for correlated traits) "results" table, which contains each of the meta-analytic means for all effect sizes of interest, for further analyses.  We further use this table as part of the Shiny App, which is able to provide the percentage differences between males and females for mean, variance and coefficient of variance.

```{r Table3}
    kable(metacombo, digits = 3) %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "200px")

#### 4. Supplemental Table S4: "Meta-analysis on Functional Trait Groups"
  SHINICHI
  This table summarizes our overall meta-analyses on the functional trait group level ("GroupingTerm"). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95% CI's, standard error and I2 values are provided.
```{r Table4}
  kable(overall2, digits = 3) %>%
   kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

#### 5. Supplemental Table S5: "Sex-biased effect sizes"
  SHINICHI
  This table provides some of the data plotted in Figures 5 and Figure S2. Here, we provide an overview over meta-analysis results performed on traits that were significantly biased to either sex. This table summarizes findings for both sexes and the respective functional trait groups.

  kable(mf_sig, digits = 3) %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "300px")
```

## Supplemental Figures
### 1. Preparation for Figure S1
#### 1.1: Preparing the data
The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)
```{r FigS1_prep}
# *Prepare data for all traits

@@

#### 1.2: Preparing heterogeneity
The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). 
As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

```{r hetero_loop, message=FALSE, warning=FALSE, include=FALSE }
# Create dataframe to store results
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n * 30), ncol = 30)))
names(results.allhetero.grouping) <- c(
  "id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR",
  "s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR",
  "s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR",
  "s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper",
  "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper", "lnRR_se"
)

@@

for (t in 1:n) {
  tryCatch(
    {
      data_par_age <- data_subset_parameterid_individual_by_age(data, t, age_min = 0, age_center = 100)

      population_stats <- calculate_population_stats(data_par_age)

      results <- create_meta_analysis_effect_sizes(population_stats)

@@

      cvr. <- metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(
        ~ 1 | strain_name, ~ 1 | production_center,
        ~ 1 | err
      ), control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), data = results)
      results.allhetero.grouping[t, 2] <- cvr.$sigma2[1]
      results.allhetero.grouping[t, 3] <- cvr.$sigma2[2]
      results.allhetero.grouping[t, 4] <- cvr.$sigma2[3]
      results.allhetero.grouping[t, 5] <- cvr.$s.nlevels[1]
      results.allhetero.grouping[t, 6] <- cvr.$s.nlevels[2]
      results.allhetero.grouping[t, 7] <- cvr.$s.nlevels[3]
      results.allhetero.grouping[t, 20] <- cvr.$b
      results.allhetero.grouping[t, 21] <- cvr.$ci.lb
      results.allhetero.grouping[t, 22] <- cvr.$ci.ub
      results.allhetero.grouping[t, 23] <- cvr.$se

@@

      vr. <- metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(
        ~ 1 | strain_name, ~ 1 | production_center,
        ~ 1 | err
      ), control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), data = results)
      results.allhetero.grouping[t, 8] <- vr.$sigma2[1]
      results.allhetero.grouping[t, 9] <- vr.$sigma2[2]
      results.allhetero.grouping[t, 10] <- vr.$sigma2[3]
      results.allhetero.grouping[t, 11] <- vr.$s.nlevels[1]
      results.allhetero.grouping[t, 12] <- vr.$s.nlevels[2]
      results.allhetero.grouping[t, 13] <- vr.$s.nlevels[3]
      results.allhetero.grouping[t, 24] <- vr.$b
      results.allhetero.grouping[t, 25] <- vr.$ci.lb
      results.allhetero.grouping[t, 26] <- vr.$ci.ub
      results.allhetero.grouping[t, 27] <- vr.$se

@@

      rr. <- metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR, random = list(
        ~ 1 | strain_name, ~ 1 | production_center,
        ~ 1 | err
      ), control = list(optimizer = "optim", optmethod = "Nelder-Mead", maxit = 1000), data = results)
      results.allhetero.grouping[t, 14] <- rr.$sigma2[1]
      results.allhetero.grouping[t, 15] <- rr.$sigma2[2]
      results.allhetero.grouping[t, 16] <- rr.$sigma2[3]
      results.allhetero.grouping[t, 17] <- rr.$s.nlevels[1]
      results.allhetero.grouping[t, 18] <- rr.$s.nlevels[2]
      results.allhetero.grouping[t, 19] <- rr.$s.nlevels[3]
      results.allhetero.grouping[t, 28] <- rr.$b
      results.allhetero.grouping[t, 29] <- rr.$ci.lb
      results.allhetero.grouping[t, 30] <- rr.$ci.ub
      results.allhetero.grouping[t, 31] <- rr.$se
    },
    error = function(e) {
      cat("ERROR :", conditionMessage(e), "\n")
    }
  )
}
```

Exclude traits without variation between mouse strains
Merge data sets containing metafor results with procedure etc. names. Dealing with correlated parameters, and running of second -order meta-analysis for  heterogeneity data.

```{r hetero}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
# nrow(results.allhetero.grouping) #232

@@

HeteroS1 <-
  heterog6 %>%
  ggplot(aes(y = GroupingTerm, x = mean)) +
  geom_errorbarh(aes(
    xmin = ci.l,
    xmax = ci.h
  ),
  height = 0.1, show.legend = FALSE
  ) +
  geom_point(aes(shape = parameter),
    fill = "black",
    color = "black", size = 2.2,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    limits = c(-0.1, 1.4),
    # breaks = c(0, 0.1, 0.2),
    name = "sigma^2"
  ) +
  # geom_vline(xintercept=0,
  # color='black',
  # linetype='dashed')+
  facet_grid(
    cols = vars(parameter), rows = vars(label),
    labeller = label_wrap_gen(width = 23),
    scales = "free",
    space = "free"
  ) +
  theme_bw() +
  theme(
    strip.text.y = element_text(angle = 270, size = 10, margin = margin(t = 15, r = 15, b = 15, l = 15)),
    strip.text.x = element_text(size = 12),
    strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
    text = element_text(size = 14),
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_blank(),
    axis.line = element_line(),
    panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
    panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.title = element_blank(),
    axis.title.x = element_text(hjust = 0.5, size = 14),
    axis.title.y = element_blank()
  )
```
#### 6. Supplemental Table S6: "Heterogeneity within groups"
  SHINICHI
  This table summarizes our findings on heterogeneity due to institutions and mouse strains. These results are based on meta-analyses on sigma^2 and errors for mouse strains and centers (Institutions), following the identical workflow from above.

  kable(het, digits = 3) %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "200px")

### 2. Figure S1

Similar to Figure 4 in the main article, this figure shows the numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits (Phase 3, Figure 3).

```{r FigS1, fig.height = 8, fig.width = 7}
#### Combined Figure S1: overall Count data, Meta anlysis results, Heterogeneity

### 3. Preparation for Figure S2
#### 3.1: Preparation for Figure S2 A
This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
```{r}
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

#### 3.2: Preparation for Figure S2 B (second-order meta-analysis)
#### 3.3: Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
Preparing data for traits with CI not overlapping 0: 
create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)
```{r FigS2_B}
#Restructure data for plotting

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```
#### 3.4: Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
Restructuring data with significant male bias (for Figure S2 B, top right panel)
```{r FigS2_B_right}
overall3.male.sigS <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key = TRUE)

@@

### 4. Figure S2

This figure is an extension of sex biases, as shown in Figure 5 in the main text. Here, we include data on lnVR (middle panel), as well as athe second method of sex-differences comparison. The panels in A show differences in numbers of affected traits, in variance (lnCVR and lnVR) and means (lnRR), where there’s a significant difference between the sexes (i.e CI not overlapping zero), and where the sex bias is greater than 10% difference (regardless of significance). Panel B depicts results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.
```{r FigS2, fig.height = 10, fig.width = 8}
library(ggpubr)
FigS2b <- ggarrange(Metameta_FigS2_female.sig, Metameta_FigS2_male.sig,
  ncol = 2, nrow = 1, widths = c(1, 1.30), heights = c(1.2, 1.2)
)

@@

## R Session Information

>>>>

---
title: "Sex and Power: Sexual dimorphism in trait variability and its eco-evolutionary and statistical implications"
author: Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky, Robert  Brooks,
  Will Cornwell, Daniel Falster, Malgorzata Lagisz, Jeremy Mason, Daniel Noble, Alistair
  Senior & Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  bookdown::html_document2: 
            toc: true
            code_folding: hide
            toc_depth: 6
            toc_float: yes
            number_sections: yes
            theme: flatly
  html_notebook:
    toc: yes
  word_document:
    toc: yes
    toc_depth: '4'
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
always_allow_html: yes
---

```{r packages1, echo = FALSE, eval = TRUE, messages = FALSE}
library(pacman)
pacman::p_load(readr, dplyr,metafor, devtools, purrr, tidyverse, tidyr, tibble, kableExtra, robumeta, ggpubr, ggplot2, png, grid, here, knitr)
```

# General Introduction

Our supplementary material follows the general structure described in our methods (Figure  \@ref(fig:FigG1)). Generally speaking, the first few sections describe the processes by which the data were cleaned and re-organised for analysis, and how the first order and second order meta-analsyes were conducted. Supplemental tables and figures that are referenecd within the main manuscript can be found at the end of the document or conviently referenecd quickly using the side table of contents panel.

```{r FigG1, fig.cap="Workflow of data process and meta-analysis"}

include_graphics(here("images", "workflow.png"))
```

# Set-up
## Packages

@@

```{r packages, echo = TRUE, eval = FALSE}
library(pacman)
pacman::p_load(readr, dplyr,metafor, devtools, purrr, tidyverse, tidyr, tibble, kableExtra, robumeta, ggpubr, ggplot2, png, grid, here)
```
## Functions 
Functions needed for preparing the data for meta analyses

### Sub-setting data
Create a function for sub-setting the data to choose only one data point per individual per trait: "data_subset_parameterid_individual_by_age"

@@

### Population statistics
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

### Extraction of effect sizes and sample variances
Function: "create_meta_analysis_effect_sizes"

@@

## Load & clean data

### Data loading and cleaning of the csv file

@@

# Meta-analyses
## Population as analysis unit 
(Step C, Figure 3 in main document)

### Loop through meta-analyses on all traits

@@

### Merging datasets & removal of non-converged traits

@@

## Meta-analysis: condensing non-independent traits 
(Step F in Figure 3 in main article)

### Dealing with Correlated Parameters, preparation

@@

### Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

### Meta-analyses on correlated (sub-)traits, using `robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

## Second-order meta-analysis for functional groups
(Section H in Figure 3 in main article)

### Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR
#### Preparing the data

@@

#### Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

# Visualising meta-analysis results
## Preparation for Figure 4 
#### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

#### Overall results of second order meta-analysis (Figure 4, Panel B)
#### Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

## Figure 4

@@

## Preparation for Figure 5
#### Preparing data for traits with CI not overlapping 0
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

## Figure 5
Sex biases in trait values across trait groups (Step J in Figure 3). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

# Supplemental Figures

```{r FigS1_prep, echo = FALSE, eval = TRUE}
### Preparing the data
#The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)

# *Prepare data for all traits

@@

```{r hetero_loop, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, include=FALSE, cache = TRUE}

### Preparing heterogeneity

#The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). 
#As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

# Create dataframe to store results
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n * 30), ncol = 30)))
names(results.allhetero.grouping) <- c(
  "id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR",
  "s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR",
  "s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR",
  "s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper",
  "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper", "lnRR_se"
)

@@

for (t in 1:n) {
  tryCatch(
    {
      results <- data %>% 
                  data_subset_parameterid_individual_by_age(t) %>% 
                  calculate_population_stats() %>% 
                  create_meta_analysis_effect_sizes() %>%
                  mutate(err = seq_len(n()))

@@

      cvr. <- metafor::rma.mv(yi = effect_size_CVR, 
                              V = sample_variance_CVR, 
                              random = list(~ 1 | strain_name, ~ 1 | production_center,~ 1 | err), 
                              control = list(optimizer = "optim", 
                                             optmethod = "Nelder-Mead", 
                                             maxit = 1000), 
                              data = results)

      results.allhetero.grouping[t, 2] <- cvr.$sigma2[1]
      results.allhetero.grouping[t, 3] <- cvr.$sigma2[2]
      results.allhetero.grouping[t, 4] <- cvr.$sigma2[3]
      results.allhetero.grouping[t, 5] <- cvr.$s.nlevels[1]
      results.allhetero.grouping[t, 6] <- cvr.$s.nlevels[2]
      results.allhetero.grouping[t, 7] <- cvr.$s.nlevels[3]
      results.allhetero.grouping[t, 20] <- cvr.$b
      results.allhetero.grouping[t, 21] <- cvr.$ci.lb
      results.allhetero.grouping[t, 22] <- cvr.$ci.ub
      results.allhetero.grouping[t, 23] <- cvr.$se

@@

      vr. <- metafor::rma.mv(yi = effect_size_VR, 
                             V = sample_variance_VR, 
                             random = list(~ 1 | strain_name, ~ 1 | production_center, ~ 1 | err), 
                             control = list(optimizer = "optim", 
                                            optmethod = "Nelder-Mead", 
                                            maxit = 1000), 
                             data = results)

      results.allhetero.grouping[t, 8] <- vr.$sigma2[1]
      results.allhetero.grouping[t, 9] <- vr.$sigma2[2]
      results.allhetero.grouping[t, 10] <- vr.$sigma2[3]
      results.allhetero.grouping[t, 11] <- vr.$s.nlevels[1]
      results.allhetero.grouping[t, 12] <- vr.$s.nlevels[2]
      results.allhetero.grouping[t, 13] <- vr.$s.nlevels[3]
      results.allhetero.grouping[t, 24] <- vr.$b
      results.allhetero.grouping[t, 25] <- vr.$ci.lb
      results.allhetero.grouping[t, 26] <- vr.$ci.ub
      results.allhetero.grouping[t, 27] <- vr.$se

@@

      rr. <- metafor::rma.mv(yi = effect_size_ROM, 
                             V = sample_variance_ROM, 
                             random = list(~ 1 | strain_name, ~ 1 | production_center, ~ 1 | err), 
                             control = list(optimizer = "optim", 
                                            optmethod = "Nelder-Mead", 
                                            maxit = 1000), 
                             data = results)

      results.allhetero.grouping[t, 14] <- rr.$sigma2[1]
      results.allhetero.grouping[t, 15] <- rr.$sigma2[2]
      results.allhetero.grouping[t, 16] <- rr.$sigma2[3]
      results.allhetero.grouping[t, 17] <- rr.$s.nlevels[1]
      results.allhetero.grouping[t, 18] <- rr.$s.nlevels[2]
      results.allhetero.grouping[t, 19] <- rr.$s.nlevels[3]
      results.allhetero.grouping[t, 28] <- rr.$b
      results.allhetero.grouping[t, 29] <- rr.$ci.lb
      results.allhetero.grouping[t, 30] <- rr.$ci.ub
      results.allhetero.grouping[t, 31] <- rr.$se
    },
    error = function(e) {
      cat("ERROR :", conditionMessage(e), "\n")
    }
  )
}
```

```{r hetero,eval=TRUE, echo = FALSE}
#Exclude traits without variation between mouse strains
#Merge data sets containing metafor results with procedure etc. names. Dealing with correlated parameters, and running of second -order meta-analysis for  heterogeneity data.

              results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
# nrow(results.allhetero.grouping) #232

@@

HeteroS1 <-
  heterog6 %>%
  ggplot(aes(y = GroupingTerm, x = mean)) +
  geom_errorbarh(aes(
    xmin = ci.l,
    xmax = ci.h
  ),
  height = 0.1, show.legend = FALSE
  ) +
  geom_point(aes(shape = parameter),
    fill = "black",
    color = "black", size = 2.2,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    limits = c(-0.1, 1.4),
    # breaks = c(0, 0.1, 0.2),
    name = "sigma^2"
  ) +
  # geom_vline(xintercept=0,
  # color='black',
  # linetype='dashed')+
  facet_grid(
    cols = vars(parameter), rows = vars(label),
    labeller = label_wrap_gen(width = 23),
    scales = "free",
    space = "free"
  ) +
  theme_bw() +
  theme(
    strip.text.y = element_text(angle = 270, size = 10, margin = margin(t = 15, r = 15, b = 15, l = 15)),
    strip.text.x = element_text(size = 12),
    strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
    text = element_text(size = 14),
    panel.spacing = unit(0.5, "lines"),
    panel.border = element_blank(),
    axis.line = element_line(),
    panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
    panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.title = element_blank(),
    axis.title.x = element_text(hjust = 0.5, size = 14),
    axis.title.y = element_blank()
  )
```

```{r, eval = TRUE, echo = FALSE}

### Preparation for Figure S2
#### Preparation for Figure S2 A
#This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

```{r FigS2_B, eval = TRUE, echo = FALSE}

#### 3.2: Preparation for Figure S2 B (second-order meta-analysis)
#### 3.3: Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
#Preparing data for traits with CI not overlapping 0: 
#create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```

```{r FigS2_B_right, eval = TRUE, echo = FALSE}
#### 3.4: Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
#Restructuring data with significant male bias (for Figure S2 B, top right panel)

@@

## Figure 5.1
Figure \@ref(fig:FigS1) is similar to Figure 4 in the main article.

```{r FigS1, fig.height = 8, fig.width = 7, fig.cap=" Numbers of either male (blue-green bars) or female (orange-red bars) biased traits  (Panel A) across functional groups, this time for lnCVR (left hand side), lnVR (middle) and lnRR (right hand side). Panel B shows effect sizes from separate meta-analysis for each functional group, and Panel C contains results of heterogeneity analyses. All three panels represent results evaluated across all traits (Phase 3, Figure 3)."}
#### Combined Figure S1: overall Count data, Meta anlysis results, Heterogeneity

## Figure 5.2

Figure \@ref(fig:FigS2) is an extension of sex biases, as shown in Figure 5 in the main text. Here, we include data on lnVR (middle panel), as well as athe second method of sex-differences comparison.

```{r FigS2, fig.height = 10, fig.width = 8, fig.cap=" A) Differences in numbers of affected traits, in variance (lnCVR and lnVR) and means (lnRR), where there’s a significant difference between the sexes (i.e CI not overlapping zero), and where the sex bias is greater than 10% difference (regardless of significance). Panel B depicts results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits."}
library(ggpubr)
FigS2b <- ggarrange(Metameta_FigS2_female.sig, Metameta_FigS2_male.sig,
  ncol = 2, nrow = 1, widths = c(1, 1.30), heights = c(1.2, 1.2)
)

@@

# Supplemental Tables
## Table 6.1

```{r Table1_prep, echo = FALSE, eval = TRUE, results='hide'}
  # Initial checks
  length(unique(data$parameter_name)) # 232 traits
  length(unique(data$parameter_group)) # 161 parameter groups
  length(unique(data$procedure_name)) # 26 procedure groups
  length(unique(data$biological_sample_id)) # 27147 individial mice  
  # To show a table with number of males and females per strain per production center:
     ST1 <- cbind(data %>% 
                   group_by(production_center, strain_name) %>% 
                   count(biological_sample_id, sex) %>% 
                   count(sex) %>% 
                   print(n = Inf))

```
```{r Table1, message=FALSE}

         kable(ST1, caption = "Summary of the available numbers of male and female mice from each strain and originating institution.") %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "200px")

## Table 6.2

```{r, TableS2}  
# Table detailing numbers of correlated and uncorrelated traits
kable(cbind(meta1 %>% count(parameter_group)), caption = "Trait categories (parameter_group) and the number of correlated traits within these categories. Traits were meta-analysed using robumeta") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

## Table 6.3

```{r Table3}
    kable(metacombo, digits = 3, caption = "We use this corrected (for correlated traits) results table, which contains each of the meta-analytic means for all effect sizes of interest, for further analyses.  We further use this table as part of the Shiny App, which is able to provide the percentage differences between males and females for mean, variance and coefficient of variance.") %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "200px")

## Table 6.4

```{r Table4}
  kable(overall2, digits = 3, caption = "Summary of overall meta-analyses on the functional trait group level (GroupingTerm). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95% CI's, standard error and I2 values are provided.") %>%
   kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

## Table 6.5

  kable(mf_sig, digits = 3, caption = "Provides an overview over meta-analysis results performed on traits that were significantly biased towards either sex. This table summarizes findings for both sexes and the respective functional trait groups.") %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "300px")
```

## Table 6.6

  kable(het, digits = 3, caption = "Summarizes our findings on heterogeneity due to institutions and mouse strains. These results are based on meta-analyses on sigma^2 and errors for mouse strains and centers (Institutions), following the identical workflow from above.") %>%
    kable_styling() %>%
    scroll_box(width = "100%", height = "200px")

# R Session Information
